---
title: "Crime + Grime Index Development"
author: "Justin Elszasz"
email: "justin.elszasz@baltimorecity.gov"
date: "Tuesday, January 21, 2020"
output:
  html_notebook:
    code_folding: hide
    fig_height: 5
    fig_width: 10
    toc: yes
    toc_depth: 1
editor_options: 
  chunk_output_type: inline
---

The purpose of this notebook is to create a neighborhood index that combines crime and cleanliness data. The index relies on [311 service request data](https://data.baltimorecity.gov/City-Services/311-Customer-Service-Requests/9agw-sxsr) and [Part 1 violent crime data](https://data.baltimorecity.gov/Public-Safety/BPD-Part-1-Victim-Based-Crime-Data/wsfq-mvij), both availble on Open Baltimore.

# Brief Description of The Index

The index primarily relies on counts of violent crime and service requests *per 1,000 residents*. This is done to remove the influence of population on the counts of these events. The per-capita metrics are then scaled to between 0 and 1 to build up the overall index.

All Part 1 crime types are used.

The SR types used are:

Illegal Dumping:

- HCD-Illegal Dumping
- SW-SIU Clean Up

Property Maintenance:

- HCD-Sanitation Property
- SW-Boarding
- SW-HGW
- SW-Cleaning

Right-of-Way Cleaning

- SW-Dirty Street
- SW-Dirty Alley


```{r setup, include = T, echo = T, message = FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = T, 
                      warning = F, 
                      message = F,
                      include = T,
                      fig.width = 4,
                      fig.height = 3)
```

```{r libraries}
library(tidyverse)
library(ggiteam)
library(lubridate)
library(readxl)
library(spdplyr)
library(leaflet)
library(RSocrata)
library(geojsonsf)
library(ggmap)
library(rgdal)
library(scales)
library(sf)
library(htmltools)


register_google("AIzaSyByyKUj4QaAlzNtPJPw-HkEEr7d4hTawO0")
```
 
```{r load_sr, cache = T}
sr_query <- paste0("https://data.baltimorecity.gov/resource/9agw-sxsr.json",
                "?$where=",
                "((srtype like 'SW-Boarding' OR ",
                "srtype like 'SW-Cleaning' OR ",
                "srtype like 'SW-HGW' OR ",
                "srtype like 'SW-Dirty Alley' OR ",
                "srtype like 'SW-Dirty Street' OR ",
                "srtype like 'HCD-Illegal Dumping' OR ",
                "srtype like 'HCD-SIU' OR ",
                "srtype like 'SW-SIU Clean Up' OR ",
                "srtype like 'HCD-Sanitation Property') AND ",
                "date_extract_y(createddate)==2019)")

sr <- read.socrata(url = sr_query)

sr <- sr %>% 
  mutate(
    latitude = as.numeric(latitude), 
    longitude = as.numeric(longitude),
    neighborhood = toupper(neighborhood))
```

```{r load_crime, cache = T}
crime_query <- paste0("https://data.baltimorecity.gov/resource/wsfq-mvij.json?$where=",
                "(Description like 'HOMICIDE' OR ",
                "Description like 'SHOOTING' OR ", 
                "Description like 'AGG. ASSAULT' OR ",
                "Description like 'RAPE' OR ",
                "contains(Description, 'ROBBERY')) AND ",
                "date_extract_y(crimedate)==2019")

crime <- read.socrata(crime_query)
```

```{r cityline}
cityline_url <- "https://data.baltimorecity.gov/resource/nfyc-pejx.geojson"

# surely there is a better way, and i tried, but couldn't get anything else to work.
# from geojson to sf to spatial df. geojson_sp didn't work for me.
cityline <- geojson_sf(cityline_url)
cityline <- as_Spatial(cityline)
cityline <- spTransform(cityline, CRS("+init=epsg:4326"))

```


```{r load_neighborhoods}
hoods.url <- "https://data.baltimorecity.gov/resource/h3fx-54q3.geojson"

# surely there is a better way, and i tried, but couldn't get anything else to work.
# from geojson to sf to spatial df. geojson_sp didn't work for me.
hoods <- geojson_sf(hoods.url)
hoods <- as_Spatial(hoods)
hoods <- spTransform(hoods, CRS("+init=epsg:4326"))
```

```{r}
# join acs data to hoods
hoods_pop <- read_excel("../data/raw/Neighborhood Pop.xlsx")
hoods@data <- left_join(hoods@data, hoods_pop, by = c("label" = "Name"))
```


```{r}
# bpd facilities to remove
facilities <- c(
  "300 E MADISON ST", # central booking
  "1400 E NORTH AVE", # district court
  "1900 ARGONNE DR", # northeastern
  "2200 W Cold Spring Ln", # northern
  "1000 N MOUNT ST", # western
  "5200 REISTERSTOWN RD	", # northwest
  "1600 EDISON HWY", # eastern
  "400 FONTHILL AVE", # southwest
  "0 CHERRY HILL RD", # southern
  "600 E FAYETTE ST", # central/hq
  "5700 EASTERN AVE", # southeast
  "2000 W BALTIMORE ST",  # bon secours
  "4000 DEEPWOOD RD", # loch raven va
  "300 N GAY ST", # juv booking
  "4900 EASTERN AV", # bayview
  "1800 ORLEANS ST", # jhh downtown
  "600 N WOLFE ST", # jhh downtown
  "0 S GREENE ST", # umd medical center
  "3400 N CALVERT ST" # union memorial
)

```

Population distribution of Baltimore neighborhoods.

```{r fig.height = 1, fig.width = 2}
hoods@data %>%
  ggplot(aes(ACS_2013_2017)) +
  geom_histogram() +
  theme_iteam_google_docs()
```

Tenth percentile for population:

```{r}
quantile(hoods$ACS_2013_2017, 0.1, na.rm = T)
```

# Analysis

```{r}
crime_labeled <- crime %>%
  filter(
    !is.na(latitude),
    !(location %in% facilities)
    ) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_join(st_as_sf(hoods), join = st_within, left = F)

crime_hood_counts <- crime_labeled %>%
  data.frame() %>%
  count(description, label) %>%
  spread(key = description, value = n) %>%
  replace(is.na(.), 0) %>%
  transmute(
    label = label,
    agg_assault = `AGG. ASSAULT`,
    rape = RAPE,
    homicide_shootings = HOMICIDE + SHOOTING,
    robbery = `ROBBERY - COMMERCIAL` +
      `ROBBERY - CARJACKING` +
      `ROBBERY - RESIDENCE` +
      `ROBBERY - STREET`) 
```

```{r}
sr_labeled <- sr %>%
  filter(
    !is.na(latitude)
    ) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_join(st_as_sf(hoods), join = st_within, left = F)

sr_hood_counts <- sr_labeled %>%
  data.frame() %>%
  count(srtype, label) %>%
  spread(key = srtype, value = n) %>%
  replace(is.na(.), 0) %>%
  transmute(
    label = label,
    illegal_dumping = `HCD-Illegal Dumping` + `SW-SIU Clean Up`,
    property_maintenance = `HCD-Sanitation Property` +
      `SW-Boarding` +
      `SW-Cleaning` +
      `SW-HGW`,
    right_of_way = `SW-Dirty Street` + `SW-Dirty Alley`
  )

```

```{r}
# to per 1,000 residents
to_per_capita <- function(x){1000 * x / hoods$ACS_2013_2017}

hood_metrics <- hoods

hood_metrics@data <- hood_metrics@data %>%
  select(-acres,
         -nbrdesc,
         -shape_leng,
         -shape_area,
         -color_2,
         -Decennial2010)

hood_metrics@data <- hood_metrics@data %>%
  left_join(
    crime_hood_counts,
    by = "label") %>%
  left_join(
    sr_hood_counts,
    by = "label"
  ) %>%
  mutate_at(
    vars(
      illegal_dumping,
      property_maintenance,
      right_of_way,
      homicide_shootings,
      robbery,
      agg_assault,
      rape),
    list(per_1000_res = to_per_capita))
```

Distribution of per capita rates of crime and service requests:

```{r}
hood_metrics@data %>%
  select(
    label,
    illegal_dumping_per_1000_res,
    property_maintenance_per_1000_res,
    right_of_way_per_1000_res,
    homicide_shootings_per_1000_res,
    robbery_per_1000_res,
    agg_assault_per_1000_res,
    rape_per_1000_res
  ) %>%
  pivot_longer(-label) %>%
  ggplot(aes(value)) +
  facet_wrap(~name, scales = "free") +
  geom_histogram(bins = 50) +
  theme_iteam_google_docs() +
  labs(title = "2019 Events per 1,000 Residents")
```

In all cases, outlier neighborhoods with very small populations skew data. 

Try removing neighborhoods with less than 150 residents (10% of neighborhoods). These are typically parks, industrial areas, etc.

```{r}
hood_metrics@data %>%
  filter(ACS_2013_2017 >= 150) %>%
  select(
    label,
    illegal_dumping_per_1000_res,
    property_maintenance_per_1000_res,
    right_of_way_per_1000_res,
    homicide_shootings_per_1000_res,
    robbery_per_1000_res,
    agg_assault_per_1000_res,
    rape_per_1000_res
  ) %>%
  pivot_longer(-label) %>%
  ggplot(aes(value)) +
  facet_wrap(~name, scales = "free") +
  geom_histogram(bins = 50) +
    theme_iteam_google_docs() +
  labs(title = "2019 Events per 1,000 Residents\nNeighborhoods with >= 150 Residents")
```

Going forward, removing neighborhoods with less than 150 residents.

```{r}
hood_metrics_over150 <- subset(hood_metrics, hood_metrics$ACS_2013_2017 > 150)
hood_metrics_over150 <- subset(hood_metrics_over150, !grepl("INDUSTRIAL", label, ignore.case = T))
```

Number of neighborhoods with at least one homicide or shooting:

```{r}
hood_metrics_over150@data %>%
  count(homicide_shootings >= 1)
```

Will likely assign 1 to neighborhoods with at least 1 homicide or shooting, 0 to all others. This will ensure neighborhoods with gun violence are given precedence.

# Building The Index

All parameters will be scaled to between 0 and 1 because variables are on different scales.

```{r}
# scale per capita
scale_index <- function(x){rescale(x, to = c(0,1))}

hood_metrics_over150@data <- hood_metrics_over150@data %>%
  mutate(
    homicide_shootings_index = ifelse(homicide_shootings > 0, 1, 0),
    rape_index = ifelse(rape > 0, 1, 0)) %>%
  mutate_at(
    vars(
      illegal_dumping_per_1000_res,
      property_maintenance_per_1000_res,
      right_of_way_per_1000_res,
      homicide_shootings_per_1000_res,
      robbery_per_1000_res,
      agg_assault_per_1000_res,
      rape_per_1000_res),
    list(index = scale_index)
    ) 
```

```{r}
# plot rescaled data
hood_metrics_over150@data %>%
  select(
    label,
    illegal_dumping_per_1000_res_index,
    property_maintenance_per_1000_res_index,
    right_of_way_per_1000_res_index,
    homicide_shootings_per_1000_res_index,
    robbery_per_1000_res_index,
    agg_assault_per_1000_res_index,
    rape_per_1000_res_index
  ) %>%
  pivot_longer(-label) %>%
  ggplot(aes(value)) +
  facet_wrap(~name, scales = "free") +
  geom_histogram(bins = 50) +
    theme_iteam_google_docs() +
  labs(title = "Standardized (between 0 and 1) 2019 Events per 1,000 Residents\nNeighborhoods with >= 150 Residents")
```

**Again, all metrics have been scaled to between 0 and 1.** Combine crime metrics into a homicide and shooting parameter (0 if none in neighborhood, 1 if at least 1) and an "other violent crime" parameter where:

$$ HomicideShootingsIndex = \begin{cases}
1 & \text{if $N_{homicides}+N_{shootings} > 0$ } \\ 
0 & \text{otherwise} \end{cases}$$

$$ OtherViolentCrimeIndex = \frac{1}{3}(RapeIndex+ AggAssaultIndex + RobberyIndex) $$

Combining the two crime indices:

$$ CrimeIndex = \frac{1}{2} (HomicideShootingsIndex + OtherViolentCrimeIndex) $$
```{r}
# create combined index
hood_metrics_over150@data <- hood_metrics_over150@data %>%
  rowwise() %>%
  mutate(
    other_violent_crime_index = 
      mean(
        c(rape_per_1000_res_index,
          agg_assault_per_1000_res_index, 
          robbery_per_1000_res_index), 
        na.rm = T),
    grime_index = mean(
      c(illegal_dumping_per_1000_res_index,
        property_maintenance_per_1000_res_index,
        right_of_way_per_1000_res_index),
      na.rm = T)
  ) %>%
  ungroup() %>%
  mutate(crime_index = 0.5 * homicide_shootings_per_1000_res_index + 
           0.5 * other_violent_crime_index,
         crime_grime_index = (2/3) * crime_index + (1/3) * grime_index,
         crime_grime_index_quadrature = sqrt(crime_index^2 + grime_index^2))
```

Now creating the "grime" index:

$$ GrimeIndex = \frac{1}{3}(IllegalDumpingIndex + PropMaintenanceIndex + RightOfWayIndex) $$
Combine into a single index. Crime weighted more heavily.

$$ CrimeGrimeIndex = \frac{2}{3}CrimeIndex + \frac{1}{3}GrimeIndex $$

# Just Show Me The Results

```{r fig.height = 8, fig.width = 8}
# base_map <- get_map(
#   location=c(lon = -76.61, lat = 39.3), 
#   zoom=12, 
#   maptype = 'toner-lite', 
#   source = 'stamen'
#   )
# 
# ggmap(base_map) +
#   # geom_point(
#   #   data = repeat_offender_locations %>%
#   #     filter(!is.na(lon),
#   #            n >= 3),
#   #   aes(x = lon, y = lat),
#   #   color = "blue",
#   #   size = 2
#   # )  
#   geom_polygon(
#     data = fortify(hood_metrics_over150) %>%
#       filter(!is.na(crime_grime_index)),
#     aes(x = long, y = lat, group = group, fill = crime_grime_index), 
#     alpha = 0.9,color = "black"
#   )# +
#   # scale_fill_viridis_c(na.value = NA, 
#   #                      #values = rescale(c(0, 20, 70), c(0,1)),
#   #                      option = "plasma",
#   #                      guide_colorbar(title = "Annual DV Incidents\nper 1,000 Residents"))
#   # 
```

```{r}
hoods@data <- hoods@data %>%
  left_join(
    hood_metrics_over150@data %>% select(-ACS_2013_2017),
    by = "label"
  )
```


```{r}
# color palette for map
pal <-  colorNumeric("viridis", 
                     domain = hoods$crime_grime_index,
                     na.color = NA)

# labels
hoods@data <- hoods@data %>%
  mutate(map_label = paste0(
    "<b>",hoods$label, "</b><br>",
    "Population: ", round(hoods$ACS_2013_2017, 0), "<br>",
    "Crime & Grime Index: ", round(hoods$crime_grime_index, 2), "<br>",
    "Crime Index: ", round(hoods$crime_index, 2), "<br>",
    "Grime Index: ", round(hoods$grime_index, 2), "<br>",
    "<br><b>2019 Crime</b><br>",
    "Homicides + Shootings: ", hoods$homicide_shootings, "<br>",
    "Rape: ", hoods$rape, "<br>",
    "Aggravated Assault: ", hoods$agg_assault, "<br>",
    "Robbery: ", hoods$robbery, "<br>",
    "<br><b>2019 Service Requests</b><br>",
    "Illegal Dumping SRs: ", hoods$illegal_dumping, "<br>",
    "Property Maintenance SRs: ", hoods$property_maintenance, "<br>",
    "Right-of-Way SRs: ", hoods$right_of_way, "<br>"
))
```

```{r fig.width = 8, fig.height = 6}
leaflet() %>%
  setView(lng = -76.6, lat = 39.3, zoom = 11) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addPolygons(data = hoods,
              weight = 2,
              fillColor  = ~pal(hoods$crime_grime_index),
              color = "black",
              opacity = 0.9,
              fillOpacity = 0.75,
              label = ~lapply(hoods$map_label, HTML)) %>%
  addControl("Baltimore<br>Crime & Grime Index", position = "topleft")
```

## Distribution of Indexes

Because of the homicide and shooting index being 1 for neighborhoods with at least one homicide or shooting and 0 for all others, the distribution is bifurcated into neighborhoods with and without shootings.

```{r, fig.height = 1, fig.width = 2}
hoods@data %>%
  ggplot(aes(crime_grime_index)) +
  geom_histogram() +
  theme_iteam_google_docs() +
  geom_text(aes(0.3, 25, 
                label="At Least 1 Homicide or Shooting\nin 2019"), size = 2, hjust = 0) +
  geom_text(aes(0.04, 25, 
                label="No Homicides or Shootings\nin 2019"), size = 2, hjust = 0)
```

## The List, In Order of Descending Crime-Grime Index

Some neighborhoods with very small populations still sneak onto the list because we only used 150 as the cut-off, but those can be manually dropped if not interested.

```{r}
the_list <- hoods@data %>%
  arrange(desc(crime_grime_index)) %>%
  select(-acres, -nbrdesc, -shape_leng, -shape_area, -color_2, -Decennial2010) %>%
  rename(neighborhood = label,
         population_acs_2013_2017 = ACS_2013_2017)

write_csv(the_list, "../data/processed/crime_grime_index_v3.csv")

the_list
```

```{r}
top_20 <- hoods %>%
  arrange(desc(crime_grime_index)) %>%
  top_n(n = 20, wt = crime_grime_index)
```

```{r fig.width = 8, fig.height = 6}
leaflet() %>%
  setView(lng = -76.6, lat = 39.3, zoom = 11) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(data = cityline,
              weight = 2,
              #fillColor  = iteam.colors[1],
              color = iteam.colors[2],
              opacity = 0.9,
              fillOpacity = 0.0) %>%
  addPolygons(data = top_20,
              weight = 2,
              fillColor  = iteam.colors[1],
              color = iteam.colors[2],
              opacity = 0.9,
              fillOpacity = 0.75,
              popup = ~lapply(top_20$map_label, HTML),
              label = top_20$label,
              labelOptions = labelOptions(noHide = T, 
                                          textOnly = T,
                                          direction = 'center',
                                          textsize = 2)) %>%

  addControl("Baltimore<br>Crime & Grime Index<br>Top 20 Neighborhoods<br><br>Click Neighborhood<br>For More Info", 
             position = "topleft")
```


