---
title: "stuff"
author: "Justin Elszasz"
email: "justin.elszasz@baltimorecity.gov"
date: "Tuesday, August 6, 2019"
output:
  html_notebook:
    code_folding: hide
    fig_height: 5
    fig_width: 10
    toc: yes
    toc_depth: 2
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = T, echo = T, message = FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = T, 
                      warning = F, 
                      message = F,
                      include = T,
                      fig.width = 10,
                      fig.height = 5)
```

```{r libraries}
library(tidyverse)
library(ggiteam)
library(lubridate)
library(readxl)
library(rgdal)
library(spdplyr)
library(leaflet)
library(RSocrata)
```

```{r load_data, echo = F}
quads <- readOGR("../data/raw/DPW Quadrant Shapefile/SW_Quadrants_Polygon_New.shp", verbose = F)

query <- paste0("https://data.baltimorecity.gov/resource/9agw-sxsr.json",
                "?$where=",
                "((srtype like 'SW-Boarding' OR ",
                "srtype like 'SW-Cleaning' OR ",
                "srtype like 'SW-HGW' OR ",
                "srtype like 'SW-Dirty Alley' OR ",
                "srtype like 'SW-Dirty Street') AND (date_trunc_y(CreatedDate) = '2019'))")
                #"(date_trunc_y(CreatedDate) = '2018')))")

sr <- read.socrata(url = query)

sr <- sr %>% mutate(latitude = as.numeric(latitude), longitude = as.numeric(longitude))

sr <- read_csv("../data/raw/ob/311_Customer_Service_Requests_2019-11-20.csv")

```

```{r}
sr_geo@data <- sr_geo@data %>%
  mutate(ClosedDate_date = mdy_hms(CloseDate),
         CreatedDate_date = mdy_hms(CreatedDate),
         DueDate_date = mdy_hms(DueDate))

sr_geo@data <- sr_geo@data %>%
  mutate(closedate_date = as.Date(closedate),
         createdate_date = as.Date(createddate),
         duedate_date = as.Date(duedate))
```


```{r}
sr_geo <- SpatialPointsDataFrame(
  coords = sr %>% 
    select(Longitude, Latitude) %>% 
    mutate_all(funs(as.numeric)) %>%
    filter(!is.na(Latitude)), 
  data = sr %>% filter(!is.na(Latitude)), 
  proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
)

sr_geo <- SpatialPointsDataFrame(
  coords = sr %>% 
    select(longitude, latitude) %>% 
    mutate_all(funs(as.numeric)) %>%
    filter(!is.na(latitude)), 
  data = sr %>% filter(!is.na(latitude)), 
  proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
)

quads <- spTransform(
  quads, 
  CRS( "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
  )
```

```{r}
dpw_facilities <- c("417 E FAYETTE ST",
                    "3939 REISTERSTOWN RD",
                    "3411 BANK ST",
                    "5225 YORK RD")
```

```{r}
sr_geo@data <- sr_geo@data %>%
  bind_cols(over(sr_geo, quads))

```

```{r fig.width = 1.6, fig.height = 1.}
sr_geo@data %>%
  mutate(year.created = year(createddate),
         od = ifelse(duedate < "2019-08-30", T, F)) %>%
  filter(year.created == 2019,
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus),
         !grepl("HCD", srtype),
         !(srtype %in% c("SW-Dirty Street", "SW-Dirty Alley"))) %>%
  count(ID, srtype) %>%
  #group_by(ID) %>%
  #mutate(quad.total = sum(n)) %>%
  ggplot(aes(ID, n)) + 
  coord_flip() +
  geom_col(aes(fill = srtype), position = position_stack(reverse = T)) +
  geom_text(aes(x = ID, y = n, 
                label = ifelse(n > 500, n, ""), 
                family = "oswald"),
            position = position_stack(reverse = F,
                                      vjust = 0.5),
            size = 2.5) +
  scale_fill_discrete_iteam() +
    theme_iteam_google_docs() + 
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.title = element_blank(),
                panel.grid.minor.x = element_line(size = .3, color = "gray90"),
        panel.grid.major.x = element_line(size = .3, color = "gray90"),
        panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        plot.caption = element_text(face = "plain",
                                    size = 7)) +
  labs(title = "Service Requests Created in 2019",
       caption = paste0("Open Baltimore as of ", today() -1)) 
```


```{r}
sr_geo@data %>%
  mutate(year.created = year(createddate),
         od = ifelse(duedate < "2019-08-30", T, F)) %>%
  filter(year.created == 2019,
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus),
         !grepl("HCD", srtype),
         (srtype %in% c("SW-Dirty Street", "SW-Dirty Alley"))) %>%
  count(ID) %>%
  mutate(pct = n / sum(n))
```



```{r fig.height = 1.2, fig.width = 3.5}
quad.pct.od <- sr_geo@data %>% 
  filter(srstatus %in% c("New", "Open", "In Progress", "Pending"),
         !grepl("HCD", srtype),
         !is.na(ID)) %>%
  mutate(od = ifelse(duedate < "2019-09-01", T, F)) %>%
  count(ID, srtype, od) %>%
  group_by(ID, srtype) %>%
  mutate(tot = sum(n)) %>%
  ungroup() %>%
  #filter(tot >= 30) %>%
  group_by(ID, srtype) %>%
  mutate(pct = n / sum(n))

quad.pct.od %>%
  filter(od == T) %>%
  select(-od, -n) %>%
  spread(key = srtype, value = pct)

quad.pct.od %>%
  filter(od == T,
         !(srtype %in% c("SW-Dirty Street", "SW-Dirty Alley"))) %>%
  ggplot(aes(ID, n)) +
  geom_col(fill = iteam.colors[1]) +
  coord_flip() +
  facet_grid( ~ srtype, scales = "free") +
  geom_text(aes(x = ID, 
                y =  n, 
                label = paste0(n, " (", round(100 *pct, 0),"%)"),
                hjust = case_when(
                  (srtype == "SW-HGW") & (n < 1500) ~ -0.1,
                  (srtype == "SW-Cleaning") & (n < 400) ~ -0.1,
                  (srtype == "SW-Boarding") & (n < 30) ~ -0.1,
                  (srtype == "SW-Dirty Alley") & (n < 2000) ~ -0.1,
                  (srtype == "SW-Dirty Street") & (n < 500) ~ -0.1,
                  TRUE ~ 1.1
                ),
                family = "oswald"), 
            size = 2.5,
            color = "gray40") +
  # scale_y_continuous(breaks = seq(0, 3500, 1000),
  #                    #limits = c(0, 4100),
  #                    minor_breaks = seq(0, 3800, 500)) +
  theme_iteam_google_docs() + 
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.minor.x = element_line(size = .3, color = "gray90"),
        panel.grid.major.x = element_line(size = .3, color = "gray90"),
        panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom",
        plot.caption = element_text(face = "plain",
                                    size = 7)) +
  labs(title = "Number of Open & Overdue SRs",
       caption = paste0("Open Baltimore as of ", today() -2)) 

```

```{r}

```



```{r fig.height = 1., fig.width = 3}
#```{r fig.width = 3.8, fig.height = .9}
quad.wo.count <- wo_geo@data %>%
  mutate(year.created = year(`Date Create`)) %>%
  filter(year.created == 2019,
         !is.na(`Clean Size`),
         !grepl("N/A", `Clean Size`),
         !is.na(ID),
         !is.na(`Clean Type`)) %>%
  count(ID, `Clean Type`, `Clean Size`) %>%
  group_by(ID, `Clean Type`) %>%
  mutate(pct = n / sum(n),
         #labels = paste0(100 * round(pct, 2), "%\n(", n, ")"))
         labels = paste0("(", n, ")"))

quad.wo.count %>%
  ggplot(aes(ID,  pct, fill = `Clean Size`)) +
  facet_grid(cols = vars(`Clean Type`)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(y = pct, label = ifelse(pct >= 0.10, n, ""), family = "oswald"),
            position=position_stack(vjust=0.5),
            size = 2.5) +
  theme_iteam_google_docs() +
  scale_fill_discrete_iteam() +
  scale_y_continuous(labels = scales::percent,
                     breaks = seq(0, 1, 0.5)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom") 
```

```{r fig.width = 8ß}
sr_geo@data %>%
  filter(year(createddate) == 2019,
         !is.na(closedate),
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus)) %>% 
  mutate(Overdue = ifelse(closedate > duedate, T, F)) %>%
  count(ID, srtype, Overdue) %>%
  group_by(ID, srtype) %>%
  mutate(pct = n / sum(n)) %>%
  ggplot(aes(srtype, pct, fill = Overdue)) +
  facet_grid(~ID) +
  coord_flip() +
  geom_col() +
  geom_text(aes(y = pct, label = n, family = "oswald"),
            position=position_stack(vjust=0.5),
            size = 2.5) +
  theme_iteam_presentations() +
  scale_fill_discrete_iteam() +
  scale_y_continuous(labels = scales::percent,  breaks = seq(0, 1, 0.5)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.spacing = unit(1, "lines"),
        legend.position = "bottom") 
```

```{r fig.height= = 8, fig.width= 8}

sr_2019 <- sr_geo@data %>%
  filter(
         !is.na(closedate),
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus)) %>%
  mutate(Overdue = ifelse(closedate > duedate, T, F))

colors <- iteam.colors[c(1,2)]

pal <- colorFactor(colors, sr_2019$Overdue)

leaflet() %>%
  setView(lng = -76.6, lat = 39.3, zoom = 12) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addCircleMarkers(data = sr_2019,
                   color = ~pal(sr_2019$Overdue),
                   radius = 2) %>%
  addPolygons(data = quads,
              fillOpacity = 0,
              color = "black",
              weight = 3,
              opacity = 1) %>%
  addLegend(colors = colors, 
            labels = c("Met", "Not Met"),
            position = "bottomleft")
```

```{r}

sr_2019 %>%
  filter(srstatus == "Closed",
         srtype == "SW-Cleaning") %>%
  mutate(closedate_date = as.Date(closedate)) %>%
  count(srtype, ID, month(closedate_date), Overdue) %>%
  group_by(srtype, ID, `month(closedate_date)`) %>%
  mutate(pct = n / sum(n)) %>%
  filter(Overdue == T) %>%
  ggplot(aes(`month(closedate_date)`, pct, color = ID)) +
  #facet_wrap(~srtype) +
  geom_line() +
  theme_iteam_presentations() +
  scale_color_discrete_iteam()

```

```{r}
sr_geo@data %>%
  count(year(ClosedDate_date), ID) %>%
  spread(key = ID, value = n)
```


```{r fig.width = 4, fig.height = 1.5}
sr_geo@data %>%
  filter(srtype %in% c(
    # "SW-Dirty Street",
    #                    "SW-Dirty Alley"),
                       "SW-Cleaning",
                       "SW-Boarding",
                       "SW-HGW"),
         !is.na(closedate_date),
         closedate_date >= "2019-07-01",
         closedate_date < "2019-11-01",
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus)) %>%
  mutate(Overdue = ifelse(closedate_date > duedate_date, T, F)) %>%
  mutate(closed_month = floor_date(closedate_date, "month")) %>%
  count(ID, srtype, closed_month, Overdue) %>%
  group_by(srtype, closed_month, ID) %>%
  mutate(pct = n / sum(n),
         tot = sum(n)) %>% 
  filter(Overdue == F,
         tot >= 20) %>%
  ggplot(aes(closed_month, pct, color = ID)) +
  facet_wrap(. ~ srtype) +
  geom_line() +
  theme_iteam_google_docs() +
  labs(x = "Month Closed",
       y = "% On-Time",
       title = "Percent On-Time by Quadrant") +
  scale_y_continuous(labels = scales::percent) +
  scale_color_discrete_iteam()
```

```{r fig.width = 4, fig.height = 2}
sr_geo@data %>%
  filter(srtype %in% c("SW-Dirty Street",
                       "SW-Dirty Alley",
                       "SW-Cleaning",
                       "SW-Boarding",
                       "SW-HGW"),
         !is.na(closedate_date),
         closedate_date >= "2019-07-01",
         closedate_date < "2019-11-01",
         !is.na(ID),
         !grepl("Duplicate", srstatus),
         !grepl("Transferred", srstatus)) %>%
  mutate(Overdue = ifelse(closedate_date > duedate_date, T, F)) %>%
  mutate(closed_month = floor_date(closedate_date, "month")) %>%
  count(ID, srtype, closed_month) %>%
  ggplot(aes(x=closed_month, y=n, color = ID, group = ID)) +
  facet_wrap(. ~ srtype, scales = "free_y") +
  geom_line() +
  theme_iteam_google_docs() +
  labs(x = "Month Closed",
       y = "SRs Closed",
       title = "SRs Closed by Quadrant") +
  scale_color_discrete_iteam()
```

